Polynomial expansions of finite volume data in a cut cell context

### 2023-04-07 example/DomainDecomposition/EBCellMerge/hoeb_geometry_test/_doc/paper.tex
progress on ebcm paper

\title{Polynomial expansions of finite volume data in a cut cell context}
\author{P. Schwartz     \footnotemark[1]
   \and N. Overton-Katz \footnotemark[1]
   \and T. Ligocki      \footnotemark[1]
   \and H. Johansen     \footnotemark[1]
   \and D. T. Graves    \footnotemark[1]
   \and D. Devendran    \footnotemark[1]
   \and O. Anteparra    \footnotemark[1]}

\maketitle
\footnotetext[1]{Lawrence Berkeley National Laboratory, Berkeley,
  CA. Research at LBNL was supported financially by the Office of
  Advanced Scientific Computing Research of the US Department of
  Energy under contract number DE-AC02-05CH11231.}

\section
\begin{abstract}

  We discuss the mechanics of forming stable local polynomial
  expansions of finite volume data in the context of cut cell grids.
  We show that significant solvability issues can arise from certain weighting
  functions.    A very rudimentary but stable cell merger algorithm is
  presented in an attempt to show that these solvability issues are
  really an artifact of small-cell instabilties.

@@

  function 
We present an algorithm to produce the necessary geometric information
for finite volume calculations in the context of Cartesian grids with
embedded boundaries.      Given an order of accuracy for the
overall calculation, we show what accuracy is required for each of the
geometric quantites and we demonstrate how to calculate the moments
using the divergence theorem.   We demonstrate that, for  a known
flux, these moments can be used to create a flux divergence of the
expected order.

\end{abstract}
\section{Introduction}

Throughout this paper, we use the following compact ``multi-index'' notation:
\begin{align*}
(\xbold - \xbar)^\pbold &= \prod\limits^D_{d=1} (x_d - {\bar x}_d)^{p_d} \\
\pbold! &= \prod\limits^D_{d = 1} p_d!
\end{align*}
Given a point in space $\xbar$, and a $D$-dimensional integer vector
$\pbold$, we define $m_v^\pbold(\xbar)$ to be the $\pbold^{th}$
moment of the volume $V$ relative to the point $\xbar$.
\begin{equation}
m_v^\pbold(\xbar)  =  \int\limits_{V} (\xbold - \xbar)^\pbold dV
\label{eqn::volmoment}
\end{equation}
Clearly, the volume of the cut cell $|V| = m_v^\zbold$,
  where $\zbold$ is the zero vector.
We define the face moments $m_{d\pm}^\pbold(\xbar)$ to be the $\pbold^{th}$  
moments (relative to the point $\xbar$) of the faces $A_{d\pm}$.
\begin{equation}
m_{d\pm}^\pbold(\xbar)  = \int\limits_{A_{d\pm}} (\xbold - \xbar)^\pbold dA
\label{eqn::facemoment}
\end{equation}
We define two moments corresponding to the embedded boundary face $A_B$: $m^{\pbold}_{B}$ and $m^{\pbold}_{B,d}$: 
\begin{equation}
m^{\pbold}_{B}(\xbar)  = \int\limits_{A_B} (\xbold
-\xbar)^\pbold \,  dA
\label{eqn::ebmomentB}
\end{equation}
and
\begin{equation}
m^{\pbold}_{B,d}(\xbar)  = \int\limits_{A_B} (\xbold
-\xbar)^\pbold \, n_d(\xbold) \, dA.
\label{eqn::ebmoment}
\end{equation}
Note that (\ref{eqn::ebmoment}) includes the normal to the embedded boundary face.

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a multi-dimensional Taylor expansion:
\begin{equation}
\psi(\xbold) = \sum_{|\qbold| < Q} \frac{1}{\qbold!}\psi^{(\qbold)}(\xbar)
\, (\xbold - \xbar)^\qbold + O(\dx^{Q}),
\label{eqn::taylor}
\end{equation}
with the multi-index partial derivative notation
\begin{equation}
\psi^{(\qbold)} = 
  \partial^{\qbold} \psi =
  \frac{\partial^{q_1}}{\partial x_1^{q_1}} \dots
  \frac{\partial^{q_D}}{\partial x_D^{q_D}} \psi \, .
\end{equation}
We express averages over volumes as 
\begin{equation*}
  \left< \nabla \cdot \Fbold \right>_V \equiv \frac{1}{|V|} 
  \int\limits_{V} \left( \nabla \cdot \Fbold \right) \, dV
\end{equation*}
We define the volume fraction $\kappa$ to be fraction of the volume
of the cell inside the solution domain, so that 
\begin{equation}
\kappa = h^{-D} |V| = h^{-D} m_v^\zbold \quad .
\label{eqn::kappa}
\end{equation}

>>>>

\title{Polynomial expansions of finite volume data in a cut cell context}
\author{N. Overton-Katz \footnotemark[1]
   \and H. Johansen     \footnotemark[1]
   \and D. T. Graves    \footnotemark[1]
   \and D. Devendran    \footnotemark[1]
   \and O. Anteparra    \footnotemark[1]}

\maketitle
\footnotetext[1]{Lawrence Berkeley National Laboratory, Berkeley,
  CA. Research at LBNL was supported financially by the Office of
  Advanced Scientific Computing Research of the US Department of
  Energy under contract number DE-AC02-05CH11231.}
\renewcommand{\thefootnote}{\fnsymbol{footnote}}\

\begin{abstract}

  We discuss the mechanics of forming stable local polynomial
  expansions of finite volume data in the context of cut cell grids.
  We show that significant solvability issues can arise from certain weighting
  functions.    A very rudimentary but stable cell merger algorithm is
  presented as we present evidence that these solvability issues may be
  an artifact of small-cell instabilties.

@@

\section{Introduction}

Embedded boundary (EB) grids are formed when one passes an  surface
through a Cartesian mesh.    For sufficiently complex geomtries, these methods
are very attactive because grid generation is a solved problem
and not computationally overwhelming \cite{trebotich-graves}.

In the current context, we form the cutting surface as the zero
surface of a function  of space $I(\xbold)$, $\xbold \in R^D$.
For smooth ($I$),  moments can be generated to any
accuracy \cite{geometry_paper}.

EB grids can produce cells that arbitrarily small.    This produces
algorithmic difficulties that generations of researchers have labored
to overcome with great success.   EB methods are used for high speed
compressible flows \cite{compresible_ns_paper, ebamrhscl} and for
slower, incompressible applications \cite{miller-trebotich}.

Formally, the underlying description of space
is given by rectangular control volumes on a Cartesian mesh
$\Upsilon_\ibold = [(\ibold-\half {\ubold})h, (\ibold+\half
{\ubold})h], \ibold \in \bigzbold^D$, where $D$ is the dimensionality
of the problem, $h$ is the mesh spacing, and ${\ubold}$ is the vector
whose entries are all one (note we use bold font $\ubold = (u_1, \dots, u_d,
\dots, u_D)$ to indicate a vector quantity).
Given an irregular domain $\Omega$, we
obtain control volumes $V_\ibold = \Upsilon_\ibold \bigcap \Omega$ and
faces $A_{\ibold,d\pm} = A_{\ibold \pm \half \ebold_d}$ which are the
intersection of the boundary of $\partial V_\ibold$ with the
coordinate planes $\{{\xbold}:x_d = (i_d \pm \half)h \}$ ($\ebold_d$ is
the unit vector in the $d$ direction).  We also
define $A_{B,\ibold}$ to be the intersection of the boundary of the
irregular domain with the Cartesian control volume: $A_{B,\ibold}
= \partial \Omega \bigcap \Upsilon_\ibold$.

\section {Finite volume notation}

Throughout this paper, we use the following compact notation:
\begin{align*}
(\xbold - \xbar)^\pbold &= \prod\limits^D_{d=1} (x_d - {\bar x}_d)^{p_d} \\
\pbold! &= \prod\limits^D_{d = 1} p_d!
\end{align*}
Given a point in space $\xbar$, and a $D$-dimensional integer vector
$\pbold$, we define $m_v^\pbold(\xbar)$ to be the $\pbold^{th}$
moment of the volume $V$ relative to the point $\xbar$.
\begin{equation}
m_v^\pbold(\xbar)  =  \int\limits_{V} (\xbold - \xbar)^\pbold dV
\label{eqn::volmoment}
\end{equation}

Given a sufficiently smooth function $\psi$, we can approximate $\psi$
in the neighborhood of $\xbar$ using a Taylor expansion to order $P$:
\begin{equation}
\psi(\xbold)  =  \sum\limits_{p < P} C^p (\xbar - \xbar)^p \label{eqn::taylor}
\end{equation}
where $C^p$ this appropriate Taylor coefficient.  In three dimensions,
\begin{equation}
  C^p =\frac(1){p!}
      \frac{\partial^{p_0}}{\partial x_0}
      \frac{\partial^{p_1}}{\partial x_1}
      \frac{\partial^{p_2}}{\partial x_2}  (\psi).
\end{equation}

In finite volume  methods, we define grid data to be
averages over volumes.   So if we discretize the smooth function
$\psi$, the average over the volume $\ibold$ is given by
\begin{equation*}
 V_\ibold <\psi>_\ibold = \int\limits_{V_\ibold} \psi(\xbold) dV
\end{equation*}
If we insert the Taylor expansion \ref{eqn::taylor}, we get a discrete
approximation to the smooth function $\psi$ that is accurate to order $P$.
\begin{equation*}
 V_\ibold <\psi>_\ibold = \sum\limits_{p < P} C^p m^p. 
\end{equation*}
This forms a local polynomial expansion of $\psi$ around the volume
$\ibold$ expressed in terms  moments ($m$)  which are the natural
products of grid generation.
Modern, higher order embedded boundary methods use this description
directly to generate stencils
 \cite{nate-stokes-paper,  dharshi-poisson, geometry-paper}.    At the core
of these algorithms is a similar procedure for finding the Taylor
coefficients $C$ which is
worth discussing.

\section
\end{document}
